### 2014-04-16 paper/conclusion.tex
submission to UIST

While today's 3D printers are not yet able to fabricate active components in-place, we suggest that 3D printed interactive devices can be created with redirection of active input and output via pipes.  This amplifies and extends the utility of existing sensing and actuation approaches.  In addition to describing the design space of pipes, we also presented \systemnamenospace, a design tool for the creation of tubes inside arbitrary 3D models, and discussed many techniques that combine nicely with pipes.  We fabricated several example objects using our tool to explore new points in the design space.

We are heartened by our success: it opens new interactive device prototyping opportunities both for makers with consumer-grade 3D printers and research and industrial labs with higher-end machines.  We are excited to release our tool into the wild and see what people build.

>>>>

Pipe mediation opens new interactive device prototyping opportunities both for makers with consumer-grade 3D printers and research and industrial labs with higher-end machines.  While today's 3D printers are not yet able to fabricate active components in-place, we suggest that 3D printed interactive devices can be created with redirection of active input and output via pipes.  This amplifies and extends the utility of existing sensing and actuation approaches.  In addition to describing the design space of pipes, we also presented \systemnamenospace, a design tool for the creation of tubes inside arbitrary 3D models, and discussed many techniques that combine nicely with pipes.  We fabricated several example objects using our tool to explore new points in the design space.  We look forward to extending our software contribution into future multi-material 3D modelling tools.

### 2014-04-16 paper/discussion.tex
submission to UIST

We are also eager to test our tool with users and get feedback from makers who use it in the wild. However, based on our own experiences, we believe our system could be readily utilized by makers and researchers, and serve as a useful tool for rapdily proptotyping interactive 3D objects.

>>>>

We are also eager to test our tool with users and get feedback from makers who use it in the wild. Based on our own experiences, we believe our system could be readily utilized by makers and researchers, and serve as a useful tool for rapdily proptotyping interactive 3D objects.

### 2014-04-16 paper/exampleobjects.tex
submission to UIST

\section{Totally Tubular Example Objects}
To evaluate and highlight our tool's capabilities, we fabricated a set of six prototype objects designed using \systemnamenospace.  To show that our approach is printer-independent, we fabricated some on a Makerbot Replicator 2 and the rest on an Objet Connex 260.  Each prototype object explores the expansion of an existing sensing or actuation technique to the surface of 3D printed objects.
%All prototypes were fabricated in a single piece, unless described otherwise. \bjoern{2 of 5 will have been fabricated otherwise -  Radio, Neon sign - both in pieces.}

@@

With inspiration from PneUIs \cite{Yao-pneui} for using gases to both provide sensation to the user through openings and deform a model internally, we created a rabbit with a pair of tubes that can simulate breathing (see Figure \ref{fig:examples}b).  When the rabbit inhales through its nose, its abdomen rises, and as it exhales its abdomen falls.  For this, we used a combination air/vacuum pump (ROB-10398): one terminal creates a vacuum while the other creates positive pressure.  
The rabbit model, the well-known Stanford Bunny\footnote{\url{http://en.wikipedia.org/wiki/Stanford_bunny}}, was downloaded from Thingiverse. We then used \systemname to add two pipes: one open pipe exiting at its mouth, and one semi-closed pipe capped at its abdomen.  We connected one pipe to each of our pump's terminals, and using a programmable power supply we mimic a rabbit's breathing pattern.  This example was printed on the Objet using flexible material (100\% TangoBlackPlus); support was flushed post-print.

@@

The radio shell was modelled in a commercial CAD tool. We included slots on the 3D model matching the shape of the electronic components that were to be embedded. The model was then imported into \systemnamenospace, where we created a network of 8 individual open pipes (3 per potentiometer, 1 for the LED, 1 for the speaker) connecting the components to the radio's base. We fabricated the radio body on our Makerbot, support-free but cut in two pieces to allow the Arduino and battery pack to be inserted. We then threaded the pipes with wires to connect the components to the microcontroller in the base.

@@

The maze is based on an SVG file we manually created and processed with our internal path routing tool.  We printed the maze on the Objet using a transparent material (100\% VeroClear) for the top and a black material (100\% TangoBlackPlus) for the bottom. The final artifact was fabricated in two halves fastened together via glue, to allow removal of support material with a high-pressure water tool.  We plugged the ends of the pipes post-print with additional material to ensure the ball does not escape.

@@

Neon art is traditionally made from hand-formed glass tubes containing neon gas.  The tubes light up when a current is passed through them.  For this type of art, the path of the tubes is of crucial importance, as it determines how the sign will look.  We designed a custom neon robot head which can be animated to ``talk''. 
The external geometry of the model was imported into \systemnamenospace, which created the associated pipes using its internal path routing tool.  We added additional endpoint-constrained tubes as exits out the rear of the model to allow us to hide the controls (see Figure \ref{fig:neon}b).  We then printed the model on the Objet using transparent material. The model was printed in two halves, cutting through the plane of the internal path, to facilitate assembly.

>>>>

\section{Totally Tubular Example Objects}
To evaluate and highlight our tool's capabilities, we fabricated a set of six prototype objects designed using \systemnamenospace.  To show that our approach is printer-independent, we fabricated some on a Makerbot Replicator 2 and the rest on a Stratasys Connex 260.  Each prototype object explores the expansion of an existing sensing or actuation technique to the surface of 3D printed objects.
%All prototypes were fabricated in a single piece, unless described otherwise. \bjoern{2 of 5 will have been fabricated otherwise -  Radio, Neon sign - both in pieces.}

@@

With inspiration from PneUIs \cite{Yao-pneui} for using gases to both provide sensation to the user through openings and deform a model internally, we created a rabbit with a pair of tubes that can simulate breathing (see Figure \ref{fig:examples}b).  When the rabbit inhales through its nose, its abdomen rises, and as it exhales its abdomen falls.  For this, we used a combination air/vacuum pump (ROB-10398): one terminal creates a vacuum while the other creates positive pressure.  
The rabbit model, the well-known Stanford Bunny\footnote{\url{http://en.wikipedia.org/wiki/Stanford_bunny}}, was downloaded from Thingiverse. We then used \systemname to add two pipes: one open pipe exiting at its mouth, and one semi-closed pipe capped at its abdomen.  We connected one pipe to each of our pump's terminals, and using a programmable power supply we mimic a rabbit's breathing pattern.  This example was printed on the Connex using flexible material (100\% TangoBlackPlus); support was flushed post-print.

@@

The radio shell was modelled in a commercial CAD tool. We included slots on the 3D model matching the shape of the electronic components that were to be embedded. The model was then imported into \systemnamenospace, where we created a network of 8 individual open pipes (3 per potentiometer, 1 for the LED, 1 for the speaker) connecting the components to the radio's base. We fabricated the radio body on our Makerbot, support-free but cut in two pieces to allow the Arduino and battery pack to be inserted. We then threaded the pipes with wires to connect the components to the microcontroller contained in the base.

@@

The maze is based on an SVG file we manually created and processed with our internal path routing tool.  We printed the maze on the Connex using a transparent material (100\% VeroClear) for the top and a black material (100\% TangoBlackPlus) for the bottom. The final artifact was fabricated in two halves fastened together via glue, to allow removal of support material with a high-pressure water tool.  We plugged the ends of the pipes post-print with additional material to ensure the ball does not escape.

@@

Neon art is traditionally made from hand-formed glass tubes containing neon gas.  The tubes light up when a current is passed through them.  For this type of art, the path of the tubes is of crucial importance, as it determines how the sign will look.  We designed a custom neon robot head which can be animated to ``talk''. 
The external geometry of the model was imported into \systemnamenospace, which created the associated pipes using its internal path routing tool.  We added additional endpoint-constrained tubes as exits out the rear of the model to allow us to hide the controls (see Figure \ref{fig:neon}b).  We then printed the model on the Connex using transparent material. The model was printed in two halves, cutting through the plane of the internal path, to facilitate assembly.

### 2014-04-16 paper/relatedwork.tex
submission to UIST

Previous work has begun to investigate how to fabricate interactive objects using particular sensing and display technologies. Printed Optics \cite{Willis-printedoptics} uses light pipes to create integrated displays and senses input optically through components added at print time, while Slyper et al., fabricate flexible robot armatures capable of sensing interaction via changes in air pressure \cite{Slyper-pressure}. Our work is inspired by these projects, and contributes a general design tool for modelling objects that can have such capabilities.

@@

@@

Some prior work has also focused on modifying interior parts of 3D objects. 
%Takayama, et al., built a system described in \cite{Takayama-fruit} which allowed users to create internal color maps of the interior of objects, allowing them to see digital ``slices'' of their ``oranges.'' \bjoern{don't understand this} Our work focuses on 3D printable geometry rather than rendering.  
In Make it Stand~ \cite{Prevost-makeitstand}, material is removed from the interior of objects with the goal of changing their center of gravity. Infrastructs~\cite{Willis-infrastructs} encodes information using interior cavities that can be sensed using terahertz imaging. Bickel, et al.,~\cite{Bickel-deformation} explored enforcing deformation behavior with multi-material blending inside objects. Modeling internal chambers is one aspect of our work, however our focus is on interaction rather than balance, identification, or deformation.  
%B\"{a}cher, et al., built an algorithm for automatically placing 3D fabricatable joints into meshes \cite{Bacher-articulated}. All these contributions are orthogonal to ours, as our focus is on precisely \emph{removing} material from printed objects, rather than blending materials or modifying components.

\subsubsection{Interaction for Fabrication vs. Fabrication for Interaction}
HCI researchers are also developing novel interaction techniques for providing input to digital fabrication machines~\cite{Mueller-constructable,Willis-interactive,mixfab}, and new hybrid hand tools that combine manual processes with digital fabrication~\cite{zoran2013freed,rivers2012position}. These projects are complementary as they target different aspects of the design process, though it is conceivable that our pipe design tools could be added to augmented reality modeling interfaces like MixFab~\cite{mixfab}.

>>>>

Previous work has begun to investigate how to fabricate interactive objects using particular sensing and display technologies. Printed Optics \cite{Willis-printedoptics} uses light pipes to create integrated displays and senses input optically through components added at print time, while Slyper et al.,~\cite{Slyper-pressure} fabricate flexible robot armatures capable of sensing interaction via changes in air pressure. Our work is inspired by these projects, and contributes a general design tool for modelling objects that can have such capabilities.

@@

\subsubsection{Interaction for Fabrication vs. Fabrication for Interaction}
HCI researchers are also developing novel interaction techniques for providing input to digital fabrication machines~\cite{Mueller-constructable,mixfab,Willis-interactive}, and new hybrid hand tools that combine manual processes with digital fabrication~\cite{rivers2012position,zoran2013freed}. These projects are complementary as they target different aspects of the design process, though it is conceivable that our pipe design tools could be added to augmented reality modeling interfaces like MixFab~\cite{mixfab}.

@@

Some prior work has also focused on modifying interior parts of 3D objects. 
%Takayama, et al., built a system described in \cite{Takayama-fruit} which allowed users to create internal color maps of the interior of objects, allowing them to see digital ``slices'' of their ``oranges.'' \bjoern{don't understand this} Our work focuses on 3D printable geometry rather than rendering.  
In Make it Stand~\cite{Prevost-makeitstand}, material is removed from the interior of objects with the goal of changing their center of gravity. Infrastructs~\cite{Willis-infrastructs} encodes information using interior cavities that can be sensed using terahertz imaging. Bickel, et al.,~\cite{Bickel-deformation} explored enforcing deformation behavior with multi-material blending inside objects. Modeling internal chambers is one aspect of our work, however our focus is interaction rather than balance, identification, or deformation.  
%B\"{a}cher, et al., built an algorithm for automatically placing 3D fabricatable joints into meshes \cite{Bacher-articulated}. All these contributions are orthogonal to ours, as our focus is on precisely \emph{removing} material from printed objects, rather than blending materials or modifying components.

### 2014-04-16 paper/tubes.tex
submission to UIST

\systemname allows users to design pipes in two ways. In one, they can select exterior anchor points on their objects; the tool then creates complete point-to-point routings using A* path search and physical rod simulation (Figure \ref{fig:tool-process-exterior}). In the other, users can import vector art describing desired interior paths (Figure \ref{fig:tool-process-interior}).  \systemname then creates a single path that follows the input shape using edge graph manipulation and Euler tour generation.  We thicken these routes to create pipes.  The resultant pipes are subtracted from the original mesh using voxelization and remeshing. Our tool is implemented in C++ as an extension to Meshmixer, a consumer 3D mesh editing tool~\cite{Schmidt-meshmixer}.

@@

\subsubsection{Pipes with Multiple Endpoints}
To allow designers to create pipes with multiple endpoints (e.g., the star topologies as in Figure \ref{fig:designspace}), we use one surface position per pipe and a central fixed endpoint that the user can interactively position (Figure~\ref{fig:boat-star}).  The user can connect several pipes, hitting a key to add each in sequence and positioning them with the mouse.  The pipes route as normal.

@@

@@

For path-constrained pipes, the route of the pipe must conform to a desired shape. A maker can import a vector graphics file (such as SVG) describing the path she wants for her pipes: all lines in this input file are interpreted as desired pipe components for the final artifact.  Using sliders, she can change the pipes' radius.  A preview is provided to show the radius and location of the cut, and the maker can align her model as appropriate.

@@

To avoid the complexities of direct mesh booleans, we construct a discretized volumetric (i.e., voxel-based) representation of the original mesh and the tube mesh.  Boolean subtraction is trivial in such representations.  We specifically use narrow-band level sets~\cite{Museth04}, which represent the surface as the zero iso-contour of an
approximate distance field.
We then use marching cubes~\cite{Lorensen87} to produce the final mesh for printing.
This strategy does incur some resampling artifacts, however so does the 3D printing, and compared to the print process, the time required to voxelize at high resolution is inconsequential.
Since we have a distance field, we can easily introduce useful effects, such as the
ability to add a thin membrane ``cap'' over the end of a tube. 
This is accomplished by applying a standard ``thicken'' operation to the level set version
of the initial mesh, intersecting with a sphere placed at the tube endpoint, and then performing
a boolean union with the initial shape-minus-tube result (Figure \ref{fig:cap}).
A straightforward extension would be to add or subtract additional elements at
the tube endpoints.

@@

\section{Fabrication Techniques}
Several pragmatic concerns arise when fabricating models with hollow pipes and cavities on additive manufactuing machines. Pipes will frequently lead to overhangs (where an underlying supporting layer has a smaller footprint than the subsequent layer).
%Beyond design, the tubes must actually be printable to be useful.  
%When geometries contain overhangs, as there may be in designed pipes and cavities, 
To allow overhangs, 3D printers lay down support material that must be later removed from the model.  This removal process can be time-intensive and challenging.  The physical fabrication process used by different machines gives rise to differing strategies for avoiding support material or easing its removal.  We next discuss techniques for fused deposition modeling (FDM) machines, such as the Makerbot\footnote{\url{http://makerbot.com}} and multi-jetting machines such as Objet\footnote{\url{http://stratasys.com}}, as well as powder-based machines (e.g., ZCorp\footnote{\url{http://3dsystems.com}}) and those that use stereolithography (e.g., Form 1\footnote{\url{http://formlabs.com}}).

@@

On single-material hobbyist FDM printers, support structures are built using the same material as the model. Such support may be impossible to remove from internal cavities.  Meshmixer has a built-in tool for minimizing support for FDM prints; it re-orients the part to keep as much surface area as possible within the 45 degree constraint.  In future work we would like to explore a tool in which a user can select surfaces (pipes) that should be prioritized for support avoidance.

\subsection{Multi-Jetting}
On multi-jetting machines, bridging is not possible and stepover potential is much less, because the machine is laying down droplets of liquid that are later UV cured.  Unsupported overhangs can be approximately $14^{\circ}$ on our Objet printer.  Multi-jetting and higher-end FDM machines use a secondary support material that must be manually removed afterwards. Some support can be melted out (Projet wax material); other support can be dissolved in lye or blasted with high-pressure water. Complex pipe geometries can make water blasting impossible and other processes more labor- and time-intensive.  All black and clear models in Figure \ref{fig:teaser} required support removal post-print: the neon sign and maze were additionally cut into two pieces to give better access to the support.

>>>>

\systemname allows users to design pipes in two ways. In one, they select exterior anchor points on their objects; the tool then creates complete point-to-point routings using A* path search and physical rod simulation (Figure \ref{fig:tool-process-exterior}). In the other, users import vector art describing desired interior paths (Figure \ref{fig:tool-process-interior}).  \systemname, using edge graph manipulation and Euler tour generation, creates a single path that follows the input shape.  We thicken these routes to create pipes.  The resultant pipes are subtracted from the user's mesh via voxelization and remeshing. Our tool is implemented in C++ as an extension to Meshmixer, a consumer 3D mesh editing tool~\cite{Schmidt-meshmixer}.

@@

\subsubsection{Pipes with Multiple Endpoints}
To allow designers to create pipes with multiple endpoints (e.g., the star topologies as in Figure \ref{fig:pipespace}), we use one surface position per pipe and a central fixed endpoint that the user can interactively position (Figure~\ref{fig:boat-star}).  The user can connect several pipes, hitting a key to add each in sequence and positioning them with the mouse.  The pipes route as normal.

@@

\begin{figure}
\centering
    \includegraphics[width=3.4in]{figures/radius.png}
\caption{A user whose pipe has too small a radius (a) can adjust it with a slider (b) to be larger (c).  PipeDream's live preview shows the real size and location of the cut before it is committed to the mesh.}
\label{fig:radius}
\end{figure}

@@

For path-constrained pipes, the route of the pipe must conform to a desired shape. A maker can import a vector graphics file (such as SVG) describing the path she wants for her pipes: all lines in this input file are interpreted as desired pipe components for the final artifact.  Using sliders, she can change the pipes' radius (Figure \ref{fig:radius}).  A preview is provided to show the radius and location of the cut, and the maker can align her model as appropriate.

@@

To avoid the complexities of direct mesh booleans, we construct a discretized volumetric (i.e., voxel-based) representation of the original mesh and the tube mesh.  Boolean subtraction is trivial in such representations.  We specifically use narrow-band level sets~\cite{Museth04}, which represent the surface as the zero iso-contour of an
approximate distance field.
We then use marching cubes~\cite{Lorensen87} to produce the final mesh for printing.
This strategy does incur some resampling artifacts, however so does the 3D printing; compared to the print process, the time required to voxelize at high resolution is inconsequential.
Since we have a distance field, we can easily introduce useful effects, such as the
ability to add a thin membrane ``cap'' over the end of a tube. 
This is accomplished by applying a standard ``thicken'' operation to the level set version
of the initial mesh, intersecting with a sphere placed at the tube endpoint, and then performing
a boolean union with the initial shape-minus-tube result (Figure \ref{fig:cap}).
A straightforward extension would be to add or subtract additional elements at
the tube endpoints.

@@

\section{Fabrication Techniques}
Several pragmatic concerns arise when fabricating models with hollow pipes and cavities on additive manufactuing machines. Pipes frequently lead to overhangs (where an underlying supporting layer has a smaller footprint than the subsequent layer).
%Beyond design, the tubes must actually be printable to be useful.  
%When geometries contain overhangs, as there may be in designed pipes and cavities, 
To allow overhangs, 3D printers may lay down support material that must be later removed from the model.  The removal process can be time-intensive and challenging.  The physical fabrication process used by different machines gives rise to differing strategies for avoiding support material or easing its removal.  We next discuss techniques for fused deposition modeling (FDM) machines, such as the Makerbot\footnote{\url{http://makerbot.com}} and polyjet machines such as Stratasys Connex\footnote{\url{http://stratasys.com}}, as well as powder-based machines (e.g., ZCorp\footnote{\url{http://3dsystems.com}}) and those that use stereolithography (e.g., Form 1\footnote{\url{http://formlabs.com}}).

@@

On single-material hobbyist FDM printers, support structures are built using the same material as the model. Such support may be impossible to remove from internal cavities.  Meshmixer has a built-in tool for minimizing support for FDM prints; it re-orients the part to keep as much surface area as possible within the 45 degree constraint.  In future work, we plan to explore a tool allowing prioritzing specific surfaces (i.e., pipes) for support avoidance, or optimizing pipes to avoid overhangs.

\subsection{Polyjet}
On polyjet machines, bridging is not possible and stepover is more constrained, because the machine is laying down droplets of liquid that are later UV cured.  Unsupported overhangs can be approximately $14^{\circ}$ on our Connex printer.  Polyjet and higher-end FDM machines use a secondary support material that must be manually removed afterwards. Some support can be melted out (Projet wax material); other varieties can be dissolved in lye or blasted with high-pressure water. Complex pipe geometries can make water blasting impossible and other processes more labor- and time-intensive.  All black and clear models in Figure \ref{fig:teaser} required support removal post-print: the neon sign and maze were additionally cut into two pieces along the plane of the pipes to give better access to the support.
